---
output:
  html_document: default
  pdf_document: default
  word_document: default
---
<h2>CSE3020 DATA VISUALIZATION J-COMPONENT</h2>
<h3>Topic: COVID-19 based data visualization and analysis</h3>
<h3>Team: Soumik Kabiraj [20BCE1504], Mayank Yadav [20BCE1674]</h3>
<h3>Slot: D2</h3><br>


***
<h3>Importing necessary libraries</h3>
```{r}
library(readr) 
library(tidyverse)
library(plotly)
library(ggthemes)
library(DT)
library(scales)
library(openair) 
library(corrplot) 
library(caTools) 
library(caret)
```


<h3>Reading datasets</h3>
```{r}
confirmed <- read_csv("C:/Users/SOUMIK/Desktop/DV J-component/time_series_covid_19_confirmed.csv",col_types = cols(.default = col_double(), `Province/State` = col_character(), `Country/Region` = col_character()))

recovered <- read_csv(file = "C:/Users/SOUMIK/Desktop/DV J-component/time_series_covid_19_recovered.csv",col_types = cols(.default = col_double(),`Province/State` = col_character(), `Country/Region` = col_character()))

deaths <- read_csv(file = "C:/Users/SOUMIK/Desktop/DV J-component/time_series_covid_19_deaths.csv",col_types = cols(.default = col_double(),`Province/State` = col_character(), `Country/Region` = col_character()))

codes <- read_csv('C:/Users/SOUMIK/Desktop/DV J-component/2014_world_gdp_with_codes.csv',col_types = cols(COUNTRY = col_character(),`GDP (BILLIONS)` = col_double(),CODE = col_character()))

data <- read.csv("C:/Users/SOUMIK/Desktop/DV J-component/Conditions_Contributing_to_COVID-19_Deaths__by_State_and_Age__Provisional_2020-2022.csv")

df <- read.csv("C:/Users/SOUMIK/Desktop/DV J-component/covid_19_india.csv")

dim(confirmed)
dim(recovered)
dim(deaths)
dim(data)
dim(df)
```

```{r}
head(confirmed)
head(recovered)
head(deaths)
head(data)
head(df)
```


<h3>DATA PREPROCESSING</h3>
```{r}
colnames(confirmed)[2]="Country"
colnames(recovered)[2]="Country"
colnames(deaths)[2]="Country"
```

```{r}
confirmed <- confirmed %>%
  gather("Date", "Confirmed", -c("Province/State", "Country", "Lat", "Long")) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

recovered <- recovered %>%
  gather("Date", "Recovered", -c("Province/State", "Country", "Lat", "Long")) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

deaths <- deaths %>%
  gather("Date", "Deaths", -c("Province/State", "Country", "Lat", "Long")) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

head(confirmed)
head(recovered)
head(deaths)
```


```{r}
#Merging confirmed, recovered and deaths dataset into one dataset
ts_total <- confirmed %>%
  left_join(deaths) %>%
  left_join(recovered) %>%
  mutate(Recovered = replace_na(Recovered, replace = 0))

head(ts_total)
dim(ts_total)
```


```{r}
#We all know "Diamond Princess" and "MS Zaandam" are cruises, So we have to remove them from the data
ts_total <- ts_total %>%
  filter(Country != "Diamond Princess") %>%
  filter(Country != "MS Zaandam")
ts_total$Deaths[is.na(ts_total$Deaths)] <- 0
```


```{r}
#Grouping by country, date and incorporating new cases
cases_latest <- ts_total %>%
  group_by(Country, Date) %>%
  summarise(Confirmed  = sum(Confirmed),Recovered = sum(Recovered),Deaths = sum(Deaths)) %>%
  mutate("New Cases" = Confirmed - lag(Confirmed, 1) ) %>%
  filter(Date == max(Date))

day_latest <- max(cases_latest$Date)

head(cases_latest)
```

```{r}
#Created a dataset including total news of COVID-19
cases_total_date <- ts_total %>%
  rename(Region = Country) %>%
  group_by(Date) %>%
  summarise(Confirmed = sum(Confirmed),
            Deaths = sum(Deaths),
            Recovered = sum(Recovered)) %>%
  mutate("New_Cases" = Confirmed - lag(Confirmed, 1))

cases_total_date$New_Cases[is.na(cases_total_date$New_Cases)] <- 0 

cases_total_latest <- cases_total_date %>%
  filter(Date == max(Date))

head(cases_total_latest)
head(cases_total_date)
```


```{r}
codes <- codes %>%
  select(COUNTRY, CODE) %>%
  rename(Region = COUNTRY ,Code = CODE) %>%
  rownames_to_column("id")
head(codes)
```

```{r}
codes$id <- as.integer(codes$id)
head(codes)
```

```{r}
codes$Region <- codes$Region %>%
  str_replace(pattern = "United States", replacement = "US")
```


```{r}
#adding country codes with cases_latest dataset
cases_latest_codes <- cases_latest %>%
  left_join(codes, by = c("Country" = "Region" )) %>%
  arrange(desc(Confirmed))
head(cases_latest_codes)
```

```{r}
cases_latest_codes %>%
  select(Country,Code, Date, Confirmed, `New Cases`, Recovered, Deaths) %>%
  arrange(desc(Confirmed)) %>%
  datatable(rownames = FALSE,options = list())
```

```{r}
cases_all <- cases_total_date %>%
  select(-Confirmed, -New_Cases) %>%
  gather("Status", "Cases", -"Date")

barchart <- ggplot(data = cases_total_date, aes(x = Date)) +
  geom_bar(aes(y = Confirmed), position = "stack", stat = "identity", fill = "#ff5050") +
  geom_bar(data = cases_all, aes(y = Cases, fill = Status), position = "stack", stat = "identity") +
  scale_fill_manual(values = c("#000000", "#009900")) +
  scale_y_continuous(breaks = seq(0, 21000000, by = 1000000), labels = comma) +
  theme(panel.background = element_rect(fill = "White"),
        legend.position = "bottom",
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()) +
  ggtitle("World COVID-19 Total Cases by Day")

ggplotly(barchart) 
```

```{r}
cases_total_date <- cases_total_date %>%
  group_by(Date, Confirmed) %>%
  mutate(Mortality_rate = Deaths / Confirmed, Recovery_rate = Recovered / Confirmed) %>%ungroup()

barchart_1 <-cases_total_date %>%
  select(Date, Mortality_rate, Recovery_rate) %>%
  gather(status.ratio, ratio, -Date ) %>%
  ggplot(aes(x = Date, y = ratio, fill = status.ratio)) +
  geom_bar(stat = "identity", position = "dodge") +
    theme(panel.background = element_rect(fill = "White"),
          legend.position = "bottom",
          axis.title = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank()) +
    ggtitle("The Mortality_rate and the Recovery_rate")

ggplotly(barchart_1) %>%
layout(legend = list(orientation = 'h'))
```


```{r}
conf<- cases_total_date[,"Confirmed"]
date <- seq(from=as.Date('2020-01-22'),by=1,to = as.Date('2021-05-29'))
calendarPlot(data.frame(conf, date), pollutant = 'Confirmed', year = 2020, main = "Confirmed Cases")
```


```{r}
rec <- cases_total_date[,"Recovered"]
calendarPlot(data.frame(rec, date), pollutant = 'Recovered', year = 2020, main = "Recovered Cases", cols = "PiYG")
```

```{r}
det <- cases_total_date[,"Deaths"]
calendarPlot(data.frame(det, date), pollutant = 'Deaths', year = 2020, main = "Deaths", cols = "RdGy")
```

```{r}
New_Cases <- cases_total_date[,"New_Cases"]
calendarPlot(data.frame(New_Cases, date), pollutant = 'New_Cases', year = 2020, main = "New Cases", cols = "BrBG")
```

```{r}
cases_total_date %>%
  select(-Date) %>%
  na.omit() %>%
  cor(use = "pairwise.complete.obs") %>%
  corrplot.mixed()
```


```{r}
top_10_confirmed <- cases_latest %>%
  select(Country, Confirmed) %>%
  arrange(desc(Confirmed))

top_10_confirmed[1:10,] %>%
  ggplot(aes(x = reorder(Country,Confirmed), y = Confirmed )) +
  geom_bar(stat = "identity", fill  = "red", width = 0.8) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "", y = "", title = "Top 10 (the Most Confirmed Cases)") +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(axis.title = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 11, face = "bold"))
```

```{r}
top_10_Deaths <- cases_latest %>%
  select(Country, Deaths) %>%
  arrange(desc(Deaths))

top_10_Deaths[1:10,] %>%
  ggplot(aes(x = reorder(Country,Deaths), y = Deaths )) +
  geom_bar(stat = "identity", fill  = "blue", width = 0.8) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "", y = "", title = "Top 10 (the Most Deaths)") +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(axis.title = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 11, face = "bold"))
```

```{r}
top_10_Recovered <- cases_latest %>%
    select(Country, Recovered) %>%
  arrange(desc(Recovered))

top_10_Recovered[1:10,] %>%
  ggplot(aes(x = reorder(Country,Recovered), y = Recovered )) +
  geom_bar(stat = "identity", fill  = "green", width = 0.8) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "", y = "", title = "Top 10 (the Most Recovered)") +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(axis.title = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 11, face = "bold"))
```

```{r}
head(cases_latest)
```

```{r}
options(repr.plot.width = 40, repr.plot.height = 9)

cases_latest %>%
    select(Country, Confirmed, Recovered, Deaths) %>%
    gather(key = group_var, value = "Cases", -Country, na.rm = TRUE) %>%
    group_by(Country, group_var) %>%
    summarise(n = sum(Cases), .groups = "drop_last") %>%
    arrange(desc(n)) %>%
    group_by(group_var) %>%
    slice(1:5)%>%
  
    ggplot(aes(x = Country, y = n, fill=Country)) +
    geom_bar(stat = "identity") +
    facet_grid(~ group_var, scales = "free") +
    scale_y_continuous(labels = scales::comma) +
    geom_label(aes(label=round(n/1000000, 1)), size=2, fill="white") +
    labs(title = "Top Countries per Case Type", subtitle = "Numbers in Millions") + theme(
    axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1))
```

```{r}
top_6_affected <- ts_total %>%
  select(Country, Date, Confirmed, Deaths, Recovered) %>%
  filter(Country %in% c("US", "Spain", "Italy", "United Kingdom",
                                 "Russia","France")) %>%
  group_by(Country, Date) %>%
  summarise(Confirmed  = sum(Confirmed),
            Recovered = sum(Recovered),
            Deaths = sum(Deaths)) %>%
  mutate("New_Cases" = Confirmed - lag(Confirmed, 1),
         "Recovery_Rate" = Recovered / Confirmed ,
         "Mortality_Rate" = Deaths / Confirmed) %>%
  mutate("New_Cases" = round(New_Cases, 3),
         "Recovery_Rate" = round(Recovery_Rate, 3),
         "Mortality_Rate" = round(Mortality_Rate, 3))

top_6_affected$New_Cases[is.na(top_6_affected$New_Cases)] <- 0
top_6_affected$Recovery_Rate[is.nan(top_6_affected$Recovery_Rate)] <- 0
top_6_affected$Mortality_Rate[is.nan(top_6_affected$Mortality_Rate)] <- 0
```


```{r}
top_6_affected %>%
  ggplot(aes(x = Date, y = Mortality_Rate, fill = Country)) +
  geom_bar(stat = "identity",alpha = 0.8) +
  facet_wrap(~ Country) +
  labs(x = "", y = "Mortality Rate") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        axis.title.y = element_text(face = "bold", size = 10))
```


<h3>COMORBIDITIES</h3>
```{r}
data %>% 
  group_by(Condition.Group) %>% 
  summarise(covid_deaths = sum(COVID.19.Deaths, na.rm = TRUE)) %>% 
  mutate(relative_frequency = covid_deaths/sum(covid_deaths)) %>% 
  arrange(desc(covid_deaths))
```

```{r}
data %>%
  group_by(Age.Group) %>% 
  summarise(covid_deaths = sum(COVID.19.Deaths, na.rm = TRUE)) %>% 
  mutate(relative_frequency = covid_deaths/sum(covid_deaths)) %>%
  filter(Age.Group != 'All Ages' & Age.Group != 'Not stated' )%>%
  arrange(desc(covid_deaths))
```

```{r}
comorb_data <- data %>% 
  filter(Age.Group != 'All Ages' & Age.Group != 'Not stated' ) %>%
  filter(Condition.Group != 'COVID-19') %>%
  drop_na(COVID.19.Deaths) %>% 
  select(-c(Flag, Number.of.Mentions, Condition.Group ,ICD10_codes, Data.As.Of, Year, Month, State, Group)) %>% 
  rename(Age_Group = Age.Group, Covid_Deaths = COVID.19.Deaths, Start_Date = Start.Date, End_Date = End.Date) %>%  
  arrange(desc(Covid_Deaths))

final_comorb <- comorb_data %>% 
  group_by(Age_Group, Condition) %>% 
  summarise(Covid_Deaths = sum(Covid_Deaths, na.rm = T)) %>% 
  mutate(Relative_Frequency = Covid_Deaths/sum(Covid_Deaths)) %>% 
  arrange(desc(Covid_Deaths))
 


head(final_comorb)
```

```{r}
options(repr.plot.width=15, repr.plot.height = 10)
final_comorb %>% 
  ggplot(mapping = aes(x = Age_Group, y = Condition))+
  geom_tile(mapping = aes(fill = Covid_Deaths))
```


<h3>Visualizations and analysis of India data</h3>
```{r}
datatable <- df %>%filter(Date == max(Date))
  
datatable %>%
  select(State.UnionTerritory, Date, Confirmed, Cured, Deaths) %>%
  arrange(desc(Confirmed)) %>%
  datatable(rownames = FALSE,options = list())
```

```{r}
Top_6 <- 
  df %>%
  dplyr::filter(Date == max(Date)) %>%
  arrange(desc(Confirmed)) 
Top_6 <- Top_6[1:6,]

percent <- paste(Top_6$State.UnionTerritory ,round(100*Top_6$Confirmed / sum(Top_6$Confirmed), 2), "%")

Top_6 %>%
  ggplot(aes(x = "", y = Confirmed, fill = percent) ) +
  geom_bar(stat = "identity",  width = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Pastel1") +
  theme(panel.grid = element_blank(), panel.background = element_blank(),
       plot.title = element_text(hjust = 0.5),
       axis.text.x = element_blank()) +
  theme_economist() +
  labs(x="", y = "", title = "Confirmed Cases Ratio", fill = "") 
```

```{r}
Top_6_all <- df %>%
  filter(State.UnionTerritory == Top_6$State.UnionTerritory)

Top_6_all %>%
  select(State.UnionTerritory, Date, Confirmed, Cured, Deaths) %>%
  gather(status, cases, -c("State.UnionTerritory", "Date")) %>%
  ggplot()+
  geom_line(aes(Date, cases, group = status, color = status), lwd = 1) +
  labs(title="Confirmed/Recovered cases over time by top 6 State/Union Territory",color = "") +
  facet_wrap(~ State.UnionTerritory, scales = "free")+ theme_fivethirtyeight() 
```


<h3>LINEAR REGRESSION</h3>
```{r}
model <- lm(Deaths ~ Confirmed + Recovered + `New Cases`, data=cases_latest)
```


```{r}
test <- summary(model)
R_sq <- test$r.squared
sprintf("The R-squared value for our model is: %f",R_sq)
```

```{r}
adj_R_sq <- test$adj.r.squared
sprintf("The adjusted R-squared value for our model is: %f",adj_R_sq)
```
INFERENCE: The obtained adjusted R_squared value for our model is 0.889553. Therefore, our model is able to explain 88.95% of the variance.<br>


<h3>K-FOLD CROSS-VALIDATION</h3>
```{r}
fitControl <- trainControl(method = "repeatedcv",   
                           number = 10,     # number of folds
                           repeats = 10)    # repeated ten times
```

```{r}
model.cv <- train(Deaths ~ Confirmed + Recovered + `New Cases`, data = cases_latest, method = "lasso", trControl = fitControl, preProcess = c('scale', 'center'))
model.cv
```

INFERENCE: We have used number of folds as 10 and number of repetitions as 10. The data has been preprocessed (scaling and centre) before feeding into the model. The final R-squared value obtained was 0.9097 ie 90.97% with the optimum fractions being 0.5. <br><br>

***